Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopies

### 2024-06-24 paper.tex
many changes
introduced several sections which need to be finished. at the end, i would like this paper to be something like the Minh Cho 1998 6WM hyper Raman paper - just outlining what the spectroscopy can do but in this case, also providing some results.

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
The implementation of a nonlinear hyper-Raman spectroscopy, which first appears at third order perturbation theory, can provide a useful analogue to these techniques to understand the spectroscopy and dynamics of isotropic systems.
Singly Vibrationally Enhanced (SIVE) spectroscopy, a type of nonlinear hyper-Raman spectroscopy, is an underdeveloped four wave mixing technique which shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.
%alternate name is hyper-(S/D)FG... which is more interesting and descriptive than SIVE imo. lmk your thoughts
We derive selection rules for singly resonant SIVE (SR-SIVE) and demonstrate that SR-SIVE response extracts hyper-Raman polarizabilities ($\beta$).
The SIVE output is shown to be slightly stronger than vibrational sum frequency generation (vSFG), heavily driven by the number density probed by each process. 
The theory of SIVE is extended to account for vibronic coupling induced by two-photon absorption, i.e. doubly resonant SIVE (DR-SIVE), akin to resonant hyper-Raman spectroscopy. 
DR-SIVE is shown to be a type of resonance IR spectroscopy and is likely able to resolve vibronic coupling.
\end{abstract}

@@

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
Singly vibrationally enhanced (SIVE) is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
Non-parametric, singly resonant SIVE (SR-SIVE) spectroscopy was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
In SR-SIVE, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to induce a scattering process to provide output.
SIVE was documented long ago to have characteristics similar to spontaneous hyper-Raman scattering spectroscopy. \cite{RN352}
Hyper-Raman scattering is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965}
Compared to Raman spectroscopy, hyper-Raman scattering cross sections are often small, making it a difficult technique to successfully implement.\cite{RN515, Kelley2010} 
The different selection rules of hyper-Raman scattering, however, make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems.
In 1998, Cho et al. proposed a parametric six wave mixing technique which was the coherent analogue of hyper-Raman scattering. \cite{Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes,\cite{RN243, Cho2000_Cascade} it is preferable to investigate coherent four wave mixing analogues to hyper-Raman scattering.

While SIVE showed promise as an upconverted infrared spectroscopy method, it was quickly supplemented by a doubly vibrationally resonant method, doubly vibrationally enhanced (DOVE) spectroscopy, to investigate intra- and intermolecular vibrational coupling. \cite{RN345, RN101, Cho2000}
As a result, investigations of SIVE stagnated; to date, there are less than a dozen studies which report on SR-SIVE. \cite{RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024}
Recent reports highlight non-negligible SIVE interference in DOVE experiments as well as the suitability of SIVE pathways to interpret ultrafast relaxation dynamics in isotropic systems. \cite{Bonn2024, McDonnell2024}
These reports motivate deeper analysis on the properties of SIVE spectroscopy.

Fully coherent, ultrafast probes of vibronic coupling would yield insight into processes that control ultrafast electronic relaxation in molecular and biological systems. \cite{Bredenbeck2015, Arsenault2021}
Some years ago, Cho proposed a parametric, doubly resonant infrared/visible pathway to probe vibronic coupling in isotropic systems. \cite{Cho2001}
In this method, a single infrared pulse is resonant with a vibrational mode, and a two photon absorption event from a near infrared (NIR) or visible pulse resonant with an electronic or vibronic state is used to generate the four wave mixing signal.
These pathways are referred to herein as DR-SIVE.
Methods analogous to DR-SIVE, where two photon absorption from the infrared pulse and single photon absorption from the visible pulse is used to generate output, i.e., stimulated Raman processes, have been reported. \cite{RN301, RN120}

An experimental complexity in executing four wave mixing experiments is the demand of three separate input pulses, often all at different frequencies.
For example, in the case of DOVE, two of the three input pulses must be scanable infrared pulses. \cite{RN345} 
Only recently have methods been developed for seamless tuning and scanning of ultrafast optical parametric amplifiers across swaths of frequency space. \cite{RN162, McDonnell2024, SkyeOPA, KyleOPA}
Since SR-SIVE demands only a single resonance to be scanned, the experiment can be performed using two input pulses: one tunable, infrared OPA to scan across vibrational resonances, and a two photon absorption from the signal process of a separate OPA, or output from the oscillator which pumps the infrared OPA.
As such, laboratories experienced in sum frequency generation (SFG) spectroscopy can perform SIVE spectroscopy to understand bulk dynamics and spectroscopy using nearly the same setup.\cite{Shen1987_CPL}

Inspired by recent work which demonstrated the presence of SIVE in ultrafast DOVE experiments,\cite{McDonnell2024} we investigate the parameters which result in both SR and DR-SIVE output. \cite{Cho2000, Bonn2024}
We develop simple expressions which relate transition dipoles and hyper-Raman polarizabilties and the overall SR-SIVE output, yielding gross selection rules. 
It is shown that a SIVE active mode must be both IR and hyper-Raman active.
Since any IR active vibration is also hyper-Raman active, \cite{Andrews1978} we demonstrate that SIVE is a coherent analogue of hyper-Raman scattering spectroscopy for IR active vibrations, i.e., SIVE is an optically upconverted infrared spectroscopy.
This is akin to sum and difference frequency generation (SFG, DFG) being coherent Raman analogues for IR active vibrations. \cite{Shen90}

We investigate properties of doubly resonant SIVE (DR-SIVE) to understand how coupling between vibrational and vibronic states can be measured through DR-SIVE.
DR-SIVE is shown to demand one and two-photon absorptions between the ground and excited state.
DR-SIVE is thus the four wave mixing analogue of doubly resonant infrared-visible sum frequency generation. \cite{Shen94}

\section{Results and Discussion}
Singly resonant SIVE (SR-SIVE) spectroscopy has potential to provide deeper insight into single quantum decoherence times and molecular orientation in condensed systems.
Similarities between SR-SIVE spectroscopy, infrared spectroscopy and spontaneous hyper-Raman scattering have been noted previously. \cite{RN352, Bonn2024}
In this section, we make the connections between SIVE and hyper-Raman scattering explicit and remark on the properties of SIVE that become apparent after inspecting the selection rules.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{wmel1.png}
	\caption{Wave Mixing Energy Level (WMEL) diagrams of the relevant SR-SIVE pathways in the (a) $\vec{k}_4 = -\vec{k}_1 + \vec{k}_2 + \vec{k}_3$ and (b) $\vec{k}_4 = \vec{k}_1 + \vec{k}_2 + \vec{k}_3$ phasematching geometries. \cite{RN286, RN352}
	$\ket{g}$ is the ground state, $\ket{v}$ is an infrared active vibration and $\ket{m}, \ket{n}$ are virtual states.
	Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. 
	These diagrams can be permuted to generate additional resonant WMEL diagrams.}
	\label{fig:sivewmel}
\end{figure}

@@

\subsection{Vibrational SR-SIVE Selection Rules}
To make the connection between SR-SIVE and hyper-Raman scattering, we investigate the gross selection rules of SR-SIVE-IR.
By propagating density matrix elements in the steady state limit, the SIVE hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{i}_{v n} \mu^{j}_{nm} \mu^{k}_{mg} \mu^{l}_{gv} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{j}_{ab}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state population.
For simplicity, we consider the results for the process diagrammed in \autoref{fig:sivewmel}a, but by multiplying by (-1) and taking a complex conjugate, which arise from the bra-side transition in \autoref{fig:sivewmel}a, results for the \autoref{fig:sivewmel}b  process are obtained.
Note that $\omega_j \rightarrow -\omega_j$ for a bra side transition.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$ and is defined as $\frac{n!}{(n-m)!}$, where $n$ is the total number of input fields (3 for any four wave mixing experiment) and $m$ is the number of distinguishable input fields.\cite{RN134} 
For a SIVE experiment using two or three distinct input fields, as considered here, D = 6.
Note that $\ket{m}, \ket{n}$ are virtual states (\autoref{fig:sivewmel}).
By contracting over the virtual states to form the hyper-Raman hyperpolarizability $\beta$ (i.e., Placzek approximation), this expression can be simplified immensely.\cite{Long1970} 
Note, however, that if $\omega_2$, $\omega_3$, as depicted in \autoref{fig:sivewmel}, are nearly resonant with any state, the Placzek approximation fails. \cite{Placzek1934, Long1970}
Assuming the validity of the Placzek approximation, \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{(ijk)}_{gv} \mu^{(l)}_{gv}}{\Delta_{gv}} \rho_{gg}
\end{equation}
As such, any mode that is SIVE active must be both hyper-Raman and IR active.
However, since any IR active vibrational mode is hyper-Raman active,\cite{Cyvin1965, Andrews1978} any IR active mode is SIVE active.
This makes SIVE an optically upconverted infrared spectroscopy of isotropic samples, as its selection rules are identically those of IR spectroscopy, but its output is generally in the visible.
Unlike SFG spectroscopy, which is dependent upon both the infrared and Raman activity of a mode (through symmetry constraints, this demands that the microscopic SFG output from centrosymmetric species vanishes),\cite{Shen90, Cotton} SIVE is a nonlinear, optically upconverted spectroscopy for any infrared active mode.

To understand SR-SIVE selection rules more finely, the dipole and first hyperpolarizability operators are Taylor expanded in terms of an n$^{\text{th}}$ normal mode coordinate $Q_n$ about equilibrium: \cite{Long1970}
\begin{subequations}
	\begin{equation}
		\mu = \mu_0 + \frac{\partial \mu}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta = \beta_{0} + \frac{\partial \beta}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
Substituting into \autoref{sivebeta}, and knowing that $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2m_n\omega_{vg}}}$ under the harmonic approximation,\cite{RN230} where $m_n$ is the mass of the $Q_n$ mode, gives the SIVE-IR-I hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
		\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{48 m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \frac{\partial \beta^{(ijk)}}{\partial Q_n} {\frac{\partial \mu^{(l)}}{\partial Q_n}}  \rho_{gg}
\end{equation}
Since this expression is non-zero in the harmonic oscillator limit, SR-SIVE output is allowed for harmonic transitions. 
Similar to SFG, SR-SIVE does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any SR-SIVE process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when the Placzek approximation is valid.

A useful attribute of this selection rule is the added ability of polarization control arising from the hyperpolarizability tensor. 
Similar to polarization control in second harmonic and sum frequency generation experiments which probe different elements of the infrared transition dipole and Raman polarizability tensor,\cite{Heinz1982} probing different elements of the hyperpolarizability tensor in SR-SIVE can greatly assist in understanding ultrafast dynamics in isotropic systems. \cite{Shen90, Bonn2024}
Polarization control in IR-pump-SIVE-probe, analogous to ultrafast transient absorption spectroscopy, would be an interesting venue to assess the sensitivity of SR-SIVE to different hyper-Raman tensor elements.

\subsection{Comparison to Sum Frequency Generation}
Since the hyper-Raman process is effectively a two-photon process, a setup using only two lasers can perform the SIVE experiment by investigating $\pm \vec{k}_1 + 2\vec{k}_2$ processes.
The other commonly used two laser CMDS method is sum frequency generation, a $\chi^{(2)}$ technique, whose phasematching is dictated by $\vec{k}_1 + \vec{k}_2$.
The SFG output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
Since SFG is dependent upon transitions that transform like both rank-one and rank-two tensors, this enforces a microscopic constraint that SFG cannot yield output for centrosymmetric species under the electric dipole moment.
This follows from simple parity arguments, where inversion centers demand states possess distinct parity selection rules. \cite{RN230}
Macroscopically, inversion symmetry in the output polarization eliminates output from centrosymmetric species under the electric dipole approximation.
As a result, the output of SFG is significantly reduced relative to most third order spectroscopies because the output only depends on surface species.
However, since SIVE spectroscopy is dependent upon a hyper-Raman transition, it is useful to compare the relative output of the two processes to identify how SIVE output compares to SFG.

To motivate the application of SIVE spectroscopy, we perform a calculation to compare the output of parametric SIVE-IR (\autoref{fig:sivewmel}b) to that of vibrational SFG (vSFG) spectroscopy.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{SIVE}}} = N_{bulk}  F(\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{SIVE}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
\end{widetext}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
In this limit,
\begin{equation}
\begin{split}
		P_{ratio} &\sim 10^4 \frac{\beta E(\omega_3)}{\alpha} \sim 10
\end{split}
\end{equation}
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the SIVE output is somewhat stronger than vibrational SFG, assuming only interfacial contributions in vSFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

>>>>

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
The implementation of a nonlinear hyper-Raman spectroscopy, which first appears at third order perturbation theory, can provide a useful analogue to these techniques to understand the spectroscopy and dynamics of isotropic systems.
Hyper Difference Frequency Generation (hyper-DFG) spectroscopy, a type of nonlinear hyper-Raman spectroscopy also known as singly vibrationally enhanced spectroscopy (SIVE), is an underdeveloped four wave mixing technique which shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.
We derive selection rules for singly resonant hyper-DFG and demonstrate it provides a simpler method to extract hyper-Raman polarizabilities ($\beta$).
The hyper-DFG output is shown to be slightly stronger than vibrational difference frequency generation (vSFG), heavily driven by the number density probed by each process. 
We demonstrate the capabilities of hyper-DFG through probing the spectra and coherence dynamics of the $\nu$(CH) modes in CH$_3$CN.
The theory of hyper-DFG is extended to account for vibronic coupling induced by two-photon absorption, i.e. doubly resonant hyper-Raman (DR-HDFG), akin to resonant hyper-Raman spectroscopy. 
DR-HDFG is shown to be a type of resonance IR spectroscopy and is likely able to resolve vibronic coupling.

\end{abstract}

@@

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
Singly vibrationally enhanced (SIVE) is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
Non-parametric, hyper difference frequency generation (hyper-DFG) spectroscopy was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
Hyper-DFG has also been refered to as singly vibrationally enhanced (SIVE) spectroscopy. \cite{RN351}
In hyper-DFG, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to induce a scattering process to provide output.
Hyper-DFG was documented long ago to have characteristics similar to spontaneous hyper-Raman scattering spectroscopy. \cite{RN352}
Hyper-Raman scattering is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965}
Compared to Raman spectroscopy, hyper-Raman scattering cross sections are often small, making it a difficult technique to successfully implement.\cite{RN515, Kelley2010} 
The different selection rules of hyper-Raman scattering, however, make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems.
In 1998, Cho et al. proposed a parametric six wave mixing technique which was the coherent analogue of hyper-Raman scattering. \cite{Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes,\cite{RN243, Cho2000_Cascade} it is preferable to investigate coherent four wave mixing analogues to hyper-Raman scattering.

While hyper-DFG showed promise as an upconverted infrared spectroscopy method, it was quickly supplemented by a doubly vibrationally resonant method, doubly vibrationally enhanced (DOVE) spectroscopy, to investigate intra- and intermolecular vibrational coupling. \cite{RN345, RN101, Cho2000}
As a result, investigations of hyper-DFG stagnated; to date, there are less than a dozen studies which report on hyper-DFG. \cite{RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024}
Recent reports highlight non-negligible hyper-DFG interference in DOVE experiments as well as the suitability of hyper-DFG pathways to interpret ultrafast relaxation dynamics in isotropic systems. \cite{Bonn2024, McDonnell2024}
These reports motivate deeper analysis on the properties of hyper-DFG spectroscopy.

Fully coherent, ultrafast probes of vibronic coupling would yield insight into processes that control ultrafast electronic relaxation in molecular and biological systems. \cite{Bredenbeck2015, Arsenault2021}
Some years ago, Cho proposed a parametric, doubly resonant infrared/visible pathway to probe vibronic coupling in isotropic systems. \cite{Cho2001}
In this method, a single infrared pulse is resonant with a vibrational mode, and a two photon absorption event from a near infrared (NIR) or visible pulse resonant with an electronic or vibronic state is used to generate the four wave mixing signal.
These pathways are referred to herein as doubly resonant hyper-DFG (DR-HDFG).
Methods analogous to DR-HDFG, where two photon absorption from the infrared pulse and single photon absorption from the visible pulse is used to generate output, i.e., stimulated Raman processes, have been reported. \cite{RN301, RN120}

An experimental complexity in executing four wave mixing experiments is the demand of three separate input pulses, often all at different frequencies.
For example, in the case of DOVE, two of the three input pulses must be scanable infrared pulses. \cite{RN345} 
Only recently have methods been developed for seamless tuning and scanning of ultrafast optical parametric amplifiers across swaths of frequency space. \cite{RN162, McDonnell2024, SkyeOPA, KyleOPA}
Since HDFG demands only a single resonance to be scanned, the experiment can be performed using two input pulses: one tunable, infrared OPA to scan across vibrational resonances, and a two photon absorption from the signal process of a separate OPA, or output from the oscillator which pumps the infrared OPA.
As such, laboratories experienced in sum frequency generation (SFG) spectroscopy can perform HDFG spectroscopy to understand bulk dynamics and spectroscopy using nearly the same setup.\cite{Shen1987_CPL}

Inspired by recent work which demonstrated the presence of HDFG in ultrafast DOVE experiments,\cite{McDonnell2024} we investigate the parameters which result in both SR and DR-SIVE output. \cite{Cho2000, Bonn2024}
We develop simple expressions which relate transition dipoles and hyper-Raman polarizabilties and the overall HDFG output, yielding gross selection rules. 
It is shown that a HDFG active mode must be both IR and hyper-Raman active.
Since any IR active vibration is also hyper-Raman active, \cite{Andrews1978} we demonstrate that HDFG is a coherent analogue of hyper-Raman scattering spectroscopy for IR active vibrations, i.e., HDFG is an optically upconverted infrared spectroscopy.
This is akin to sum and difference frequency generation (SFG, DFG) being coherent Raman analogues for IR active vibrations. \cite{Shen90}

We investigate properties of DR-HDFG to understand how coupling between vibrational and vibronic states can be measured.
DR-HDFG is shown to demand one and two-photon absorptions between the ground and excited state.
DR-HDFG is thus the four wave mixing analogue of doubly resonant infrared-visible sum frequency generation. \cite{Shen94}

\section{Steady State Hyper-DFG Spectroscopy}
\subsection{Singly Resonant HDFG}
Singly resonant hyper difference generation spectroscopy (HDFG) spectroscopy has potential to provide deeper insight into single quantum decoherence times and molecular orientation in condensed systems.
Similarities between HDFG spectroscopy, infrared spectroscopy and spontaneous hyper-Raman scattering have been noted previously. \cite{RN352, Bonn2024}
In this section, we make the connections between HDFG and hyper-Raman scattering explicit and its properties.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{wmel1.png}
	\caption{Wave Mixing Energy Level (WMEL) diagrams of the relevant HDFG pathways in the (a) $\vec{k}_4 = -\vec{k}_1 + \vec{k}_2 + \vec{k}_3$ and (b) $\vec{k}_4 = \vec{k}_1 + \vec{k}_2 + \vec{k}_3$ phasematching geometries. \cite{RN286, RN352}
	$\ket{g}$ is the ground state, $\ket{v}$ is an infrared active vibration and $\ket{m}, \ket{n}$ are virtual states.
	Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. 
	These diagrams can be permuted to generate additional resonant WMEL diagrams.}
	\label{fig:sivewmel}
\end{figure}

@@

To make the connection between HDFG and hyper-Raman scattering, we investigate its gross selection rules.
By propagating density matrix elements in the steady state limit, the HDFG hyperpolarizability is \cite{RN119}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{i}_{v n} \mu^{j}_{nm} \mu^{k}_{mg} \mu^{l}_{gv} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{j}_{ab}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state population.
For simplicity, we consider the results for the process diagrammed in \autoref{fig:sivewmel}a, but by multiplying by (-1) and taking a complex conjugate, which arise from the bra-side transition in \autoref{fig:sivewmel}a, results for the \autoref{fig:sivewmel}b  process are obtained.
Note that $\omega_j \rightarrow -\omega_j$ for a bra side transition.
D, the Maker-Terhune degeneracy factor, accounts for permutation symmetry in $\gamma_{ijkl}$ and is defined as $\frac{n!}{(n-m)!}$, where $n$ is the total number of input fields (3 for any four wave mixing experiment) and $m$ is the number of distinguishable input fields.\cite{RN134} 
For a HDFG experiment using two or three distinct input fields, as considered here, D = 6.
Note that $\ket{m}, \ket{n}$ are virtual states (\autoref{fig:sivewmel}).
By contracting over the virtual states to form the hyper-Raman hyperpolarizability $\beta$ (i.e., Placzek approximation), this expression can be simplified immensely.\cite{Long1970} 
Note, however, that if $\omega_2$, $\omega_3$, as depicted in \autoref{fig:sivewmel}, are nearly resonant with any state, the Placzek approximation fails. \cite{Placzek1934, Long1970}
Assuming the validity of the Placzek approximation, \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar}\frac{\beta^{(ijk)}_{gv} \mu^{(l)}_{gv}}{\Delta_{gv}} \rho_{gg}
\end{equation}
Taylor expanding the dipole and first hyperpolarizability operators in terms of an n$^{\text{th}}$ normal mode coordinate $Q_n$ about equilibrium: \cite{Long1970}
\begin{subequations}
	\begin{equation}
		\mu = \mu_0 + \frac{\partial \mu}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta = \beta_{0} + \frac{\partial \beta}{\partial Q_n} Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
Substituting into \autoref{sivebeta}, and knowing that $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2m_n\omega_{vg}}}$ under the harmonic approximation,\cite{RN230} where $m_n$ is the mass of the $Q_n$ mode, gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{48 m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \frac{\partial \beta^{(ijk)}}{\partial Q_n} {\frac{\partial \mu^{(l)}}{\partial Q_n}}  \rho_{gg}
\end{equation}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for harmonic transitions. 
Similar to SFG, HDFG does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any HDFG or hyper sum frequency generation (HSFG) process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when the Placzek approximation is valid.

\autoref{SIVEselection} shows any HDFG active mode must be both hyper-Raman and IR active.
However, since any IR active vibrational mode is hyper-Raman active,\cite{Cyvin1965, Andrews1978} any IR active mode is HDFG active.
This makes HDFG an optically upconverted infrared spectroscopy of isotropic samples, as its selection rules are identically those of IR spectroscopy, but its output is generally in the visible.
A useful attribute of this selection rule is the added ability of polarization control arising from the hyperpolarizability tensor. 
Similar to polarization control in second harmonic and sum frequency generation experiments which probe different elements of the infrared transition dipole and Raman polarizability tensor,\cite{Heinz1982} probing different elements of the hyperpolarizability tensor in HDFG can greatly assist in understanding ultrafast dynamics in isotropic systems. \cite{Shen90, Bonn2024}
Polarization control in IR-pump-HDFG-probe, analogous to ultrafast transient absorption spectroscopy, would be an interesting venue to assess the sensitivity of HDFG to different hyper-Raman tensor elements.

Since the hyper-Raman process is stimulated by a two-photon process, a setup using only two lasers can perform the HDFG experiment by investigating $\pm \vec{k}_1 + 2\vec{k}_2$ processes, assuming a two photon absorption from a single pulse stimulates the hyper-Raman process.
The other commonly used two laser CMDS method is sum frequency generation, a $\chi^{(2)}$ technique, whose phasematching is dictated by $\vec{k}_1 + \vec{k}_2$.
The SFG output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
Since SFG is dependent upon transitions that transform like both rank-one and rank-two tensors, this enforces a microscopic constraint that SFG cannot yield output for centrosymmetric species under the electric dipole moment.
This follows from simple parity arguments, where inversion centers demand states possess distinct parity selection rules. \cite{RN230}
Macroscopically, inversion symmetry in the output polarization eliminates output from centrosymmetric species under the electric dipole approximation.
As a result, the output of SFG is significantly reduced relative to most third order spectroscopies because the output only depends on surface species.
However, since HDFG is dependent upon a hyper-Raman transition, it is useful to compare the relative output of the two processes to identify how HDFG output compares to SFG.

To motivate the application of HDFG spectroscopy, we perform a calculation to compare HDFG and vibrational SFG (vSFG) output.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{HDFG}}} = N_{bulk}  F(-\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(-\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(-\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{HDFG}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
\end{widetext}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
In this limit, $P_\text{ratio} \sim 10$.
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the HDFG output is somewhat stronger than vibrational SFG, assuming only interfacial contributions in vSFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.
